Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-08 shanghai.tex
(1) Still working on connected algebraic groups. (2) I made a home for the lemma we discussed yesterday.

  We may now lift the $\ell$-adic character $\rho_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$ determined by the structure map for $G$. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{L}_0 \iso \gcs{L}$.
  Since $b_G^*$ is full and faithful (again, see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}),
  $
  b_G^* : \Hom(m^*\cs{L}_0,\cs{L}_0\boxtimes\cs{L}_0) \to \Hom({\bar m}^*\gcs{L},\gcs{L}\boxtimes\gcs{L})
  $ 
  is a bijection. Let $\mu_0 : m^*\cs{L}_0 \to \cs{L}_0\boxtimes\cs{L}_0$ be the isomorphism matching 
  $\mu : {\bar m}^*\gcs{L} \to \gcs{L}\boxtimes\gcs{L}$, from the definition of $\cs{L}$. 
  Then $(\cs{L}_0,\mu_0)$ is an object in $\CS_0(G)$ 
  and $\cs{L} = (\gcs{L},\mu,F)$ is isomorphic to $(b_G^*\cs{L}_0,b_G^*\mu_0)$ in $\CS(G)$.
  Thus, $b_G^* : \CS_0(G) \to \CS(G)$ is essentially surjective.

@@

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$. We
  will show that the group isomorphism does not depend on these
  choices.

  First, however, we show that such global sections exist.  Since
  $G(\bFq)$ is finitely-generated abelian, it admits a decomposition
  $G(\bFq) \iso G(\bFq)_\text{free} \oplus G(\bFq)_\text{tors}$ such
  that $G(\bFq)_\text{tors}$ is finite.  Let $x_1, \ldots , x_r$ be a
  basis for the free part $G(\bFq)_\text{free}$ of $G(\bFq)$ and let
  $y_1, \ldots , y_t$ be a generators for the torsion part of
  $G(\bFq)$ corresponding to the decomposition
  $G(\bFq)_\text{tors} \iso \oplus_{j=1}^t \ZZ/m_j\ZZ$, with
  $m_j \vert m_{j+1}$. For each such $x_i$ pick $s(x_i)\in \gcs{L}_{x_i}$;
  for each such $y_j$ pick $s(y_j) \in \gcs{L}_{y_j}$; also set
  $s({\bar e}) \ceq 0 \in \gcs{L}_{\bar e}$. ....
  \Clifton{I'll finish this later.}
  In this way we define a section
  $s \ceq s_\cs{L} : G(\bFq) \to \coprod_{z\in G(\bFq)} \gcs{L}_z$ of
  projection $\coprod_{z\in G(\bFq)} \gcs{L}_z \to G(\bFq)$ such that

  \begin{equation}\label{section}
    \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
  \end{equation}
  Under the equivalence at the beginning of this section, $s = s_\cs{L}$
  corresponds to a global section of the line bundle $(\gcs{L},\mu)$.

\Clifton{I'm going to remove all mention of line bundles from this paper.}

  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $F_x : \gcs{L}_{\Frob(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $F_x(s(\Frob(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $t_\cs{L}(\Frob)(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad F_{x}(s(\Frob(x)) = t_\cs{L}(\Frob)(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $t_\cs{L}(\Frob)(x+y) = t_\cs{L}(\Frob)(x) \ t_\cs{L}(\Frob)(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

>>>>

  We may now lift the $\ell$-adic character $\rho_g : \Gal(\bFq/\Fq) \to \EEx$
  to an $\ell$-adic character $\pi_1(G,\bg) \to \EEx$ using the canonical topological group homomorphism
  $\pi_1(G,\bg) \to \Gal(\bFq/\Fq)$ determined by the structure map for $G$. But $\bg$ also
  determines an equivalence between the category of $\ell$-adic
  representations of $\pi_1(G,\bg)$ and $\ell$-adic local systems on $G$. Let
  $\cs{L}_0$ be a local system on $G$ in the isomorphism class
  determined by this $\ell$-adic character of $\pi_1(G,\bg)$.
  Then $b_G^*\cs{L}_0 \iso \gcs{L}$.
  Since $b_G^*$ is full and faithful (again, see \cite{SGA7.2}*{Expos\'e XIII} or even \cite{BBD}*{Prop. 5.2.1}),
  $
  b_G^* : \Hom(m^*\cs{L}_0,\cs{L}_0\boxtimes\cs{L}_0) \to \Hom({\bar m}^*\gcs{L},\gcs{L}\boxtimes\gcs{L})
  $ 
  is a bijection. Let $\mu_0 : m^*\cs{L}_0 \to \cs{L}_0\boxtimes\cs{L}_0$ be the isomorphism matching 
  $\mu : {\bar m}^*\gcs{L} \to \gcs{L}\boxtimes\gcs{L}$, from the definition of $\cs{L}$. 
  Then $(\cs{L}_0,\mu_0)$ is an object in $\CS_0(G)$ 
  and $\cs{L} = (\gcs{L},\mu,F)$ is isomorphic to $(b_G^*\cs{L}_0,b_G^*\mu_0)$ in $\CS(G)$.
  Thus, the full and faithful functor $b_G^* : \CS_0(G) \to \CS(G)$ is also essentially surjective,
  hence an equivalence (when $G$ is a connected, commutative algebraic group over $\Fq$).

  In Proposition~\ref{prop:bounded} we saw that $\CS_f(G)$ is a full subcategory of $\CS_0(G)$;
  more precisely, we exhibited a full and faithful functor $L_G : \CS_f(G) \to \CS_0(G)$.
  We now show that this functor is essentially surjective 
  when $G$ is a connected, commutative algebraic group over $\Fq$.
  Let $(\cs{L}_0,\mu_0)$ be a bounded character sheaf.

@@

  To produce the isomorphism between $\CSiso{G(\bFq)}$ and
  $\Hh^1(\Weil{\Fq}, G(\bFq)^*)$ we must pick a global section of the line
  bundle $(\gcs{L},\mu)$, for each character sheaf $\cs{L}$ on $G$; 
  such a section exists by Lemma~\ref{lemma:section}.
  We will show that the group isomorphism does not depend on these
  choices.
%
  We use the section $s_\cs{L}$ to define a cocycle $z_\cs{L} \in Z^1(\Weil{\Fq},G(\bFq)^*)$
  as follows. Recall that $\cs{L}$ also determines a family of linear transformations
  $F_x : \gcs{L}_{\Frob(x)} \to \gcs{L}_x$ of one-dimensional $\EE$-vector
  spaces, so $F_x(s(\Frob(x)))$ is a scalar multiply of $s(x)$, for
  each $x\in G(\bFq)$; let $t_\cs{L}(\Frob)(x)$ be that scalar, so
  \begin{equation}\label{t}
    \forall x\in G(\bFq),\qquad F_{x}(s(\Frob(x)) = t_\cs{L}(\Frob)(x) s(x).
  \end{equation}
  Conditions \ref{cs.0} and \ref{cs.2}, together with \eqref{section}, guarantee that
  the scalar $z_\cs{L}(\Frob)(x)$ is non-zero, while condition~\ref{cs.1} forces
  $t_\cs{L}(\Frob)(x+y) = t_\cs{L}(\Frob)(x) \ t_\cs{L}(\Frob)(y)$,
  for all $x,y \in G(\bFq)$.
  % \Clifton{The alternative to this approach is to change the
  %   definition of character sheaves to incorporate the notion of
  %   Galois sheaves. That might not be so bad, but would add another
  %   layer of complication, which we wouldn't use profitably, expect
  %   here.}

@@

\begin{lemma}
 If $\cs{L}$ is a character sheaf then there is a non-trivial section of $\gcs{L}$ 
 compatible with $\mu$ in the sense that 
 \begin{equation}\label{section}
  \forall x,y \in G(\bFq), \qquad \mu_{x,y}(s(x+y)) = s(x)\otimes s(y).
 \end{equation}
\end{lemma}

\begin{proof}
First, however, we show that such global sections exist.  Since
  $G(\bFq)$ is finitely-generated abelian, it admits a decomposition
  $G(\bFq) \iso G(\bFq)_\text{free} \oplus G(\bFq)_\text{tors}$ such
  that $G(\bFq)_\text{tors}$ is finite.  Let $x_1, \ldots , x_r$ be a
  basis for the free part $G(\bFq)_\text{free}$ of $G(\bFq)$ and let
  $y_1, \ldots , y_t$ be a generators for the torsion part of
  $G(\bFq)$ corresponding to the decomposition
  $G(\bFq)_\text{tors} \iso \oplus_{j=1}^t \ZZ/m_j\ZZ$, with
  $m_j \vert m_{j+1}$. For each such $x_i$ pick $s(x_i)\in \gcs{L}_{x_i}$;
  for each such $y_j$ pick $s(y_j) \in \gcs{L}_{y_j}$; also set
  $s({\bar e}) \ceq 0 \in \gcs{L}_{\bar e}$. ....
  \todo{under construction}
  In this way we define a section
  $s \ceq s_\cs{L} : G(\bFq) \to \coprod_{z\in G(\bFq)} \gcs{L}_z$ of
  projection $\coprod_{z\in G(\bFq)} \gcs{L}_z \to G(\bFq)$ such that
  $\mu_{x,y}(s(x+y)) = s(x)\otimes s(y)$ for all $x,y \in G(\bFq)$.
  Under the equivalence at the beginning of this section, $s = s_\cs{L}$
  corresponds to a global section of the line bundle $(\gcs{L},\mu)$.

\Clifton{I'm going to remove all mention of line bundles from this paper.}
\end{proof}
